A Phoneme-weighted Levenshtein Distance: A Measure of Cross-linguistic Phonological Similarity

### 2020-08-06 Communications/2020-07-30_meeting.Rmd
Prepare report for meeting

library(dplyr)
library(papaja)
library(tidyr)
library(tibble)
library(stringr)
library(forcats)
library(ggplot2)
library(here)
library(data.table)
library(tidybayes)
library(gt)
library(lubridate)

participants <- fread(here("Data", "01_participants.csv"), na.strings = "") %>% as_tibble() %>% mutate(date = as_date(date))
dat <- fread(here("Data", "01_processed_coded.csv"), na.strings = "") %>% as_tibble()
posterior <- fread(here("Results", "posterior_draws.csv"), na.strings = "") %>% as_tibble()
predictions <- fread(here("Results", "posterior_predictions.csv"), na.strings = "") %>% as_tibble()
fit <- readRDS(here("Results", "fit3_phon.rds"))
comparison <- readRDS(here("Results", "model_comparison.rds"))
item_effects <- fread(here("Results", "item_effects.csv"), na.strings = "") %>% as_tibble()
group_effects <- fread(here("Results", "group_effects.csv"), na.strings = "") %>% as_tibble()
corrs <- fread(here("Results", "corrs.csv"), na.strings = "") %>% as_tibble()

valid_participant <- filter(participants, valid_participant) %>% pull(participant)
```

@@

This resulted in `r nrow(participants)-sum(participants$valid_participant)` participants excluded, leaving `r nrow(filter(participants, valid_participant))` valid participants and `r dat %>% filter(response_type %in% c("correct", "typo", "wrong", "false_friend")) %>% nrow() %>% printnum(big.mark = ",")` valid trials.

```{r participants_group, echo=FALSE}

dat %>%
    filter(participant %in% valid_participant, response_type %in% c("correct", "typo", "wrong", "false_friend")) %>% 
    group_by(group, participant) %>% 
    summarise(n_trials = n()) %>% 
    group_by(group)%>%
    summarise(n = n(),
              n_trials = sum(n_trials)) %>% 
    gt() %>% 
    cols_label(group = "Group", n_trials = "Trials")

@@

```{r participants_language, echo=FALSE}

participants %>%
    ggplot(aes(l2, fill = l2)) +
    facet_wrap(~group, ncol = 2) +
    geom_bar(stat = "count", show.legend = FALSE) +
    labs(x = "L2", y = "Count", title = "What languages others than you native one do you know?") +
    theme_classic() +
    theme(panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
          axis.text.x = element_text(angle = 330, hjust = 0))

```{r participants_prof_l2, echo=FALSE}

participants %>%
    filter(!(l2 %in% "None")) %>% 
    pivot_longer(c(l2oral, l2written), names_to = "modality", values_to = "proficiency") %>%
    mutate(modality = str_to_sentence(str_remove(modality, "l2"))) %>% 
    ggplot(aes(l2, proficiency, fill = l2)) +
    facet_wrap(group~modality) +
    geom_col(show.legend = FALSE) +
    labs(x = "L2", y = "Self rated proficiency (0-5)") +
    theme_classic() +
    theme(panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
          axis.text.x = element_text(angle = 330, hjust = 0))

```{r participants_prof_spacat, echo=FALSE}

participants %>%
    pivot_longer(c(spanish_oral, spanish_written, catalan_oral, catalan_written), names_to = "language", values_to = "proficiency") %>%
    separate(language, c("language", "modality"), sep = "_") %>% 
    mutate(modality = str_to_sentence(str_remove(modality, "l2")),
           language = str_to_sentence(language)) %>% 
    ggplot(aes(proficiency)) +
    facet_wrap(group~language~modality, nrow = 3) +
    geom_bar(position = position_dodge(), fill = "black") +
    labs(y = "Count", x = "Self rated proficiency (0-5)") +
    scale_x_continuous(limits = c(0, 5)) +
    theme_classic() +
    theme(panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"))

@@

## Data analysis

We fit Bayesian logistic models to model the probability of participants typing a correct translation in each trial. As predictors, we included the number of more frequent phonological neighbours of the translation (`PTHN`, standardised), whether both translation equivalents started with the same phoneme (`same_onset`, with levels `different`, `same_vowel`, `same_consonant`, helmert contrast comparing different vs. same and same vowel vs. same consonant), the highest number of consecutive phonemes shared (`consecutive_longest`, standardised), and our meausure of phonological similarity (`similarity_phop`, standardised). We first for a null model `fit0_null` that only included PTHN as predictor. We then extended the model incrementally by adding other predictors. The models were finally, we fitted a model that only included `PTHN` and the traditional orthographical Levenshtein similarity (`similarity`, standardised), we test whether our measure predicted participants responses better than just relying on orthographic similarity. The models were implemente in the `brms` R package using the following formulas:

* `fit0_null <- correct | total ~ pthn + (1 + pthn | group/trial_id)`
* `fit1_onset <- correct | total ~ pthn + same_onset + (1 +  same_onset | group/trial_id)`
* `fit2_consec <- correct | total ~ pthn + same_onset + consecutive_longest + (1 + same_onset + consecutive_longest | group/trial_id)`
* `fit3_phon <- correct | total ~ pthn + same_onset + consecutive_longest + similarity_phon + (1 + same_onset + consecutive_longest + similarity_phon | group/trial_id)`
* `fit4_ort <- correct | total ~ pthn + similarity_ort + (1 + similarity_ort | group/trial_id)`

@@

@@

```{r waic, echo=FALSE}

waic <- comparison %>%
    unlist() %>%
    as.data.frame() %>% 
    rownames_to_column("model")

colnames(waic) <- c("Model", "WAIC")

waic %>%
    gt()

@@

```{r posterior_draws, echo=FALSE}

posterior %>%
    ggplot(aes(.value, .variable)) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    stat_slab(fill = "black") + 
    stat_interval(.width = c(0.95, 0.89, 0.67, 0.5), position = position_nudge(y = -0.15)) +
    labs(x = "Estimate", y = "Coefficient", colour = "Credible interval",
         title = "Posterior distribution of coefficients of fixed effects") +
    scale_colour_brewer(palette = "Blues") +
    theme_classic() +
    theme(panel.grid.major.y = element_line(colour = "grey"),
          legend.position = "top",
          legend.key = element_rect(fill = "transparent"))

@@

## Posterior predictors

```{r posterior_predictions, echo=FALSE}

predictions %>%
    ggplot(aes(similarity_phon, .value, colour = same_onset, fill = same_onset)) +
    facet_grid(consecutive_longest~pthn)+
    stat_summary(fun = mean, geom = "line", size = 1) +
    geom_vline(xintercept = 0, colour= "grey") +
    geom_hline(yintercept = 0.5, colour= "grey") +
   labs(x = "Phonological similarity", y = "Probability of correct translation",
         fill = "Credible interval", colour = "Onset",
         title = "Effect of phonological similarity on probability of correct translation",
         subtitle = "Model predictions are displayed for different quartiles of PTHN") +
    scale_fill_brewer(palette = "Set1") +
    scale_colour_brewer(palette = "Set1") +
    theme_classic() +
    theme(legend.position = "top")

@@

@@

### Item-level effects

```{r item_effects, echo=FALSE}

item_effects %>%
  pivot_wider(names_from = estimate, values_from = value) %>% 
  mutate(param = factor(param, levels = c("intercept", "sameonset1","sameonset2", "consecutivelongest", "similarityphon"),
                        labels = c("Intercept", "Different onset vs. Same onset", "Same onset: Vowel vs. Consonant", "Consecutive longest", "Phonological similarity"))) %>% 
  ggplot(aes(estimate, fill = param)) +
  facet_wrap(~param, scales = "free") +
  geom_histogram(bins = 20) +
  geom_rug(alpha = 0.1) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(x = "Estimate", y = "Density", fill = "Parameter") +
  scale_fill_brewer(palette = "Set1") +
  theme_classic() +
  theme(legend.position = "none",
        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
        panel.grid.minor.y = element_line(colour = "grey", linetype = "dotted"))

@@

### Group-level effects

```{r goup_effects_table, echo=FALSE}

group_effects %>%
  pivot_wider(names_from = estimate, values_from = value) %>% 
  mutate(param = factor(param, levels = c("intercept", "sameonset1","sameonset2", "consecutivelongest", "similarityphon"),
                        labels = c("Intercept", "Different onset vs. Same onset", "Same onset: Vowel vs. Consonant", "Consecutive longest", "Phonological similarity"))) %>% 
  mutate_if(is.numeric, round, 2) %>% 
  mutate(hdi = paste0("[", lower, ", ", upper, "]")) %>% 
  select(param, group, error, hdi) %>% 
  group_by(param) %>% 
  gt() %>% 
  cols_label(group = "Group", error = "SE", hdi = "95% HDI")

@@

```{r group_effects, echo=FALSE}

group_effects %>%
  pivot_wider(names_from = estimate, values_from = value) %>% 
  mutate(param = factor(param, levels = c("intercept", "sameonset1","sameonset2", "consecutivelongest", "similarityphon"),
                        labels = c("Intercept", "Different onset vs. Same onset", "Same onset: Vowel vs. Consonant", "Consecutive longest", "Phonological similarity"))) %>% 
  ggplot(aes(group, estimate)) +
  facet_wrap(~param) +
  geom_hline(yintercept = 0) +
  geom_errorbar(aes(ymin = lower, ymax = upper, colour =  group), width = 0, size = 5) +
  geom_errorbar(aes(ymin = estimate-error, ymax = estimate+error), width = 0, size = 1, alpha = 0.5) +
  geom_point(size = 3) +
  labs(x = "Group", y = "Estimate", fill = "Group") +
  scale_fill_brewer(palette = "Set1") +
  theme_classic() +
  theme(legend.position = "none",
        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
        panel.grid.minor.y = element_line(colour = "grey", linetype = "dotted"))

### Correlations

```{r correlations, echo=FALSE}

corrs %>% 
  filter(str_detect(variable, "trial_id")) %>%
  mutate(variable = str_remove(variable, "cor_group:trial_id__"),
         variable = str_replace(variable, "__" , " - ")) %>% 
  ggplot(aes(value, variable, xmin = lower, xmax = upper)) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = "red") +
  geom_errorbarh(show.legend = FALSE, height = 0) +
  geom_point(size = 2, show.legend = FALSE) +
  labs(x = "Estimate", y = "Correlation") +
  theme_classic() +
  theme(axis.title.y = element_blank())

>>>>

library(dplyr)
library(papaja)
library(tidyr)
library(tibble)
library(janitor)
library(stringr)
library(forcats)
library(ggplot2)
library(patchwork)
library(here)
library(data.table)
library(broom)
library(brms)
library(effectsize)
library(pairwiseComparisons)
library(tidybayes)
library(gt)
library(lubridate)

participants <- fread(here("Data", "02_participants.csv"), na.strings = c("", "NA")) %>%
  as_tibble() %>%
  mutate(date = as_date(date),
         group = as.factor(group))
contrasts(participants$group) <-  contr.helmert(3)/2

trials <- fread(here("Data", "trial_stats.csv"),
                na.strings = c("", "NA")) %>%
  as_tibble() %>%
  clean_names()

accuracy <- fread(here("Data", "03_accuracy_coded.csv"),
                  na.strings = c("", "NA")) %>%
  as_tibble() %>% 
  mutate(correct_coded = response_type %in% c("correct", "typo"),
         correct_coded = as.numeric(correct_coded))

posterior <- fread(here("Results", "posterior_draws.csv"),
                   na.strings = c("", "NA")) %>%
  as_tibble()

prior <- fread(here("Results", "prior_draws.csv"),
               na.strings = c("", "NA")) %>%
  as_tibble()

predictions <- fread(here("Results", "posterior_predictions.csv"),
                     na.strings = c("", "NA")) %>%
  as_tibble()

fit <- readRDS(here("Results", "fit4_phon.rds"))
loos <- readRDS(here("Results", "loos.rds"))
r2 <- readRDS(here("Results", "r2.rds"))

item_effects <- fread(here("Results", "item_effects.csv"),
                      na.strings = c("", "NA")) %>%
  as_tibble()

group_effects <- fread(here("Results", "group_effects.csv"),
                       na.strings = c("", "NA")) %>%
  as_tibble()

corrs <- fread(here("Results", "corrs.csv"),
               na.strings = c("", "NA")) %>%
  as_tibble()

valid_participant <- filter(participants, valid_participant) %>% pull(participant)

clean_varnames <- function(x){
  str_remove({{ x }}, "b_") %>% 
    str_replace_all(., "pthn", "PTHN") %>% 
    str_replace_all(., "similarity_phon", "Phon. Similarity") %>% 
    str_replace_all(., ":", " \U000D7 ")
}

@@

This resulted in `r nrow(participants)-sum(participants$valid_participant)` participants excluded, leaving `r nrow(filter(participants, valid_participant))` valid participants and `r accuracy %>% filter(response_type %in% c("correct", "typo", "wrong", "false_friend")) %>% nrow() %>% printnum(big.mark = ",")` valid trials.

```{r participants_prof_l2, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

participants %>%
  filter(!(l2 %in% "None")) %>% 
  pivot_longer(c(l2oral, l2written), names_to = "modality", values_to = "proficiency") %>%
  mutate(modality = str_to_sentence(str_remove(modality, "l2"))) %>%
  group_by(l2, country, modality) %>% 
  summarise(mean = mean(proficiency, na.rm = TRUE),
            n = n(), .groups = "drop") %>%
  arrange(n) %>% 
  mutate(l2 = fct_inorder(as.factor(l2))) %>% 
  ggplot(aes(l2, mean, fill = l2)) +
  facet_grid(country~modality) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(y = mean-0.5, label = paste0("n = ", n)), size = 3) +
  labs(x = "L2", y = "Self-rated proficiency (0-5)") +
  scale_x_discrete(guide = guide_axis(n.dodge = 2)) +
  scale_y_continuous(limits = c(0, 5)) +
  theme_classic() +
  theme(panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
        axis.title.x = element_blank(),
        axis.text = element_text(colour = "black"))

@@

```{r participants_prof_spacat, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

participants %>%
  pivot_longer(c(spanish_oral, spanish_written, catalan_oral, catalan_written), names_to = "language", values_to = "proficiency") %>%
  separate(language, c("language", "modality"), sep = "_") %>% 
  mutate(modality = str_to_sentence(str_remove(modality, "l2")),
         language = as.factor(str_to_sentence(language))) %>% 
  group_by(language, country, modality, proficiency) %>% 
  summarise(n = n(), .groups = "drop") %>% 
  ggplot(aes(x = proficiency, y = n, fill = country)) +
  facet_grid(language~modality) +
  geom_col() +
  labs(y = "Count", x = "Self rated proficiency (0-5)", fill = "Country") +
  scale_x_continuous(limits = c(0, 5)) +
  theme_classic() +
  theme(panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
        axis.text = element_text(colour = "black"))

@@

## Stimuli

```{r stimuli, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}

plot_onset <- trials %>%
  select(trial_id, word, language, matches("same_onset")) %>% 
  rename_all(str_remove, "same_") %>% 
  pivot_longer(matches("onset"), names_to = "language_pair", values_to = "onset") %>%  
  mutate(language_pair = str_to_upper(str_remove(language_pair, "onset_")),
         onset = ifelse(str_detect(onset, "same"), "Same", "Different")) %>% 
  drop_na(onset) %>% 
  ggplot(aes(onset, colour = language_pair, fill = language_pair)) +
  facet_wrap(~language_pair) +
  geom_bar(position = position_dodge(width = 1), na.rm = TRUE) +
  labs(x = "Onset", y = "Counts",
       colour = "Language pair", fill = "Language pair") +
  theme_classic() +
  theme(legend.position = "none")

plot_subs <- trials %>%
  select(trial_id, word, language, matches("close_")) %>% 
  rename_all(str_remove, "close_") %>% 
  pivot_longer(matches("substitutions"), names_to = "language_pair", values_to = "substitutions") %>% 
  mutate(language_pair = str_to_upper(str_remove(language_pair, "substitutions_"))) %>% 
  ggplot(aes(language_pair, substitutions, colour = language_pair, fill = language_pair)) +
  geom_violin() +
  stat_summary(fun = "mean", geom = "point", colour = "black", shape = "_", size = 10) +
  labs(x = "Close substitutions", y = "Phonological similarity",
       colour = "Language pair", fill = "Language pair") +
  scale_y_continuous(limits = c(0, 1)) +
  theme_classic() +
  theme(legend.position = "none",
        axis.title.x = element_blank())

plot_sim <- trials %>%
  select(trial_id, word, language, matches("similarity")) %>% 
  rename_all(str_remove, "similarity_") %>% 
  rename_at(vars(4:9), ~paste0("phon_", .)) %>% 
  rename_at(vars(4:9), str_replace, "phon_ort_", "ort_") %>% 
  pivot_longer(4:9, names_to = c("modality", "language_pair"), values_to = "similarity", names_sep = "_") %>% 
  pivot_wider(names_from = modality, values_from = similarity) %>% 
  mutate(language_pair = str_to_upper(language_pair)) %>% 
  ggplot(aes(ort, phon, colour = language_pair, fill = language_pair)) +
  geom_point(alpha = 0.5, size = 0.25) +
  geom_smooth(formula = "y ~ x", method = "lm") +
  labs(x = "Orthographical similarity", y = "Phonological similarity",
       colour = "Language pair", fill = "Language pair") +
  scale_y_continuous(limits = c(0, 1)) +
  scale_x_continuous(limits = c(0, 1)) +
  theme_classic()

plot_onset + plot_subs + plot_sim + guide_area() +
  plot_layout(nrow = 2, guides = "collect") &
  theme(axis.text = element_text(colour = "black"),
        panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"))

```
## Data analysis

We fit Bayesian logistic models to model the probability of participants typing a correct translation in each trial. As predictors, we included the number of more frequent phonological neighbours of the translation (`PTHN`, standardised), whether both translation equivalents started with the same phoneme (`same_onset`, with levels `different`, `same_vowel`, `same_consonant`, sum-coded as `different` (-0.5) vs. `same` (0.5), the highest number of consecutive phonemes shared (`consecutive_longest`, standardised), and our measure of phonological similarity (`similarity_phop`, standardised). We first for a null model `fit0_null` that only included PTHN as predictor. We then extended the model incrementally by adding other predictors. The models were finally, we fitted a model that only included `PTHN` and the traditional orthographical Levenshtein similarity (`similarity`, standardised), we test whether our measure predicted participants responses better than just relying on orthographic similarity. The models were implemented in the `brms` R package using the following formulas:

* `fit1_pthn <- correct | trials(total) ~ pthn + (1 + pthn | group)`
* `fit2_onset <- correct | trials(total) ~ pthn*same_onset + (1 +  pthn*same_onset | group)`
* `fit3_close <- correct | trials(total) ~ pthn*close_substitutions + (1 + pthn*close_substitutions | group)`
* `fit4_phon <- correct | trials(total) ~ pthn*similarity_phon + (1 + pthn*similarity_phon | group)`
* `fit5_ort <- correct | trials(total) ~ pthn*similarity_ort + (1 + pthn*similarity_ort | group)`

@@

```{r accuracy_anova, message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}

model <- lm(prop_correct ~ group, data = participants)
anov <- anova(model) %>% 
  tidy()

es <- anov$sumsq[1]/(anov$sumsq[1]+anov$sumsq[2])
pairs <- pairwise_comparisons(data = participants,
                              x = group,
                              y = prop_correct, type = "parametric",
                              p.adjust.method = "bonferroni") %>% 
  select(-c(label, test.details, p.value.adjustment, significance))

We performed a one-way ANOVA to analyse how participants' accuracy (the proportion of correct trials) changed across groups the three groups (ENG-SPA, ENG-CAT, and SPA-CAT). The `group` predictor explained a significant proportion of the variance, *F*(`r printnum(anov$df[1])`, `r printnum(anov$df[2])`) = `r printnum(anov$statistic[1])`, `r printp(anov$p.value[1])`, $\hat{\eta}^2$ = `r printnum(es)`). Follow-up *t*-tests revealed that Spanish-native speakers translating Catalan words were more accurate than English-native participants translating Spanish words, *t*(`r printnum(pairs$df[3])`) = `r printnum(pairs$t.value[3])`, `r printp(pairs$p.value[3])`, Cohen's *d* = `r printnum(pairs$t.value[3]/sqrt(nrow(filter(participants, group %in% c("SPA-CAT", "ENG-SPA")))))`), or Catalan words, *t*(`r printnum(pairs$df[2])`) = `r printnum(pairs$t.value[2])`, `r printp(pairs$p.value[2])`, Cohen's *d* = `r printnum(pairs$t.value[2]/sqrt(nrow(filter(participants, group %in% c("SPA-CAT", "ENG-SPA")))))`). We did not find evidence against accuracy being similar in both English-native groups, , *t*(`r printnum(pairs$df[1])`) = `r printnum(pairs$t.value[1])`, `r printp(pairs$p.value[1])`, Cohen's *d* = `r printnum(pairs$t.value[1]/sqrt(nrow(filter(participants, group %in% c("SPA-CAT", "ENG-SPA")))))`). Bonferroni corrections were applied in the aforementioned contrasts.

```{r accuracy, echo=FALSE, fig.cap="Participants' overall accuracy across groups. The X-axis represents the native language of participants (English or Spanish) and the language participants listened during the task (Spanish or Catalan). The Y-axis represents the proportion of correct. Each datapoint corresponds to one participant. Horizontal black lines represent group means."}

ggplot(participants, aes(group, prop_correct, fill = group)) +
  geom_violin(colour = NA) +
  geom_boxplot(fill = "white", width = 0.1) +
  labs(x = "Group", y = "Accuracy (% correct trials)", fill = "Group") +
  scale_y_continuous(limits = c(0, 1)) +
  theme_classic() +
  theme(panel.grid.major.y = element_line(colour = "grey", linetype = "dotted"),
        legend.position = "none",
        axis.title.x = element_blank(),
        axis.text = element_text(colour = "black"))
```

@@

```{r comparison, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}

r2_table <- r2 %>%
  mutate(r2_ci95 = paste0("[", round(q2_5, 2), ", ", round(q97_5, 2), "]")) %>% 
  rename(r2 = estimate, r2_se = est_error) %>% 
  select(model, r2, r2_se, r2_ci95)
loos$diffs %>% 
  as.data.frame() %>% 
  rownames_to_column("model") %>% 
  select(model, elpd_diff, se_diff, elpd_loo, se_elpd_loo) %>% 
  left_join(r2_table) %>% 
  relocate(model, r2, r2_se, r2_ci95) %>% 
  gt() %>% 
  fmt_number(c(2, 3, 5, 6, 7, 8)) %>% 
  cols_label(model = "Model", r2 = "R2", r2_se = "SE", r2_ci95 = "95% CrI", elpd_diff = "\U0394 ELPD", se_diff = "\U0394 SE", elpd_loo = "ELPD LOO", se_elpd_loo = "SE ELPD LOO") %>%
  tab_spanner(label = "R2", columns = matches("r2")) %>% 
  tab_spanner(label = "PSIS-LOO", columns = matches("loo|diff"))

@@

```{r posterior_coefs, echo=FALSE, message=FALSE, warning=FALSE, paged.print=TRUE}

summary(fit)$fixed %>%
  as.data.frame() %>% 
  rownames_to_column("term") %>% 
  clean_names() %>% 
  mutate(ci95 = paste0("[", round(l_95_percent_ci, 2), ", ", round(u_95_percent_ci, 2), "]"),
         term = clean_varnames(term)) %>% 
  select(-matches("percent")) %>% 
  relocate(term, estimate, est_error, ci95) %>% 
  gt(rowname_col = "term") %>% 
  fmt_number(columns = vars(estimate, est_error, rhat, bulk_ess, tail_ess)) %>% 
  cols_label(estimate = "Estimate", est_error = "SE", ci95 = "95% CrI", rhat = "R-hat", bulk_ess = "Bulk ESS", tail_ess = "Tail ESS")

@@

```{r posterior_draws, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Posterior distribution of the extended model's (fit3_phon) coefficients. The X-axis represents all the values that each coefficient can take. The Y-axis represents the probability density at each value, i.e., the probability of the coefficient taking a specific value. Grey bars colour-code highest density intervals (HDI) of different width. Each interval indicates the narrowest interval that can be define within each distribution capturing a given proportion of the probability mass (e.g., given our data there is a 95% probability that the true value of a coefficient lies within the 95% HDI interval."}

posterior %>% 
  mutate(.variable = clean_varnames(.variable)) %>% 
  ggplot(aes(.value, .variable)) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  stat_slab(show.legend = FALSE, fill = "grey30") + 
  stat_interval(.width = c(0.95, 0.89, 0.67, 0.5), position = position_nudge(y = -0.2)) +
  labs(x = "Estimate", y = "Coefficient", colour = "Credible interval",
       title = "Posterior distribution of coefficients of fixed effects") +
  scale_colour_brewer(palette = "Oranges") +
  theme_classic() +
  theme(panel.grid.major.y = element_line(colour = "grey"),
        legend.position = "top",
        legend.key = element_rect(fill = "transparent", colour = NA),
        axis.text = element_text(colour = "black"))

@@

## Posterior predictions

```{r posterior_predictions, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Predicted probability of correct translation across different measures of phonological similarity scores. Predictions were generated by sampling 1000 random draws from the posterior distribution of each predictor in the extended model (fit3_phon)."}

predictions %>%
  ggplot(aes(similarity_phon, .value)) +
  geom_point(data = accuracy, aes(y = correct_coded), shape = "|") +
  geom_line(aes(group = .draw), alpha = 0.25, colour = "grey") +
  stat_summary(fun = "mean", geom = "line", colour = "blue") +
  labs(x = "Phonological similarity", y = "Probability of correct translation",
       fill = "PTHN", colour = "Group") +
  theme_classic() +
  theme(axis.text = element_text(colour = "black"))
```

@@

```{r group_effects, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Group-level effects. We specified a maximal random effects structure. As a result, intercepts and slopes were estimated for each group. This figure shows how the estimated intercepts and slopes varied across groups. There are not any remarkable differences across groups, except for the slope of phonological similarity. The accuracy of Spanish-native speakers translating from Catala \\(SPA-CAT\\) was slightly boosted by the phonological similarity between translation equivalents in comparison with the rest of the groups \\(ENG-SPA, ENG-CAT\\)."}

group_effects %>%
  mutate(param = clean_varnames(param)) %>% 
  ggplot(aes(.value, group)) +
  facet_wrap(~param, scales = "free") +
  stat_interval(.width = c(0.95, 0.89, 0.67, 0.50)) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  labs(x = "Estimate", y = "Group", colour = "HDI (%)") +
  scale_colour_brewer() +
  theme_classic() +
  theme(legend.position = "top",
        panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
        axis.title.y = element_blank(),
        axis.text = element_text(colour = "black"))

@@

```{r correlations, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE, fig.cap="Pairwise correlations between group-level effects. This figure displaya"}

corrs %>% 
  mutate(.variable = str_remove(.variable, "cor_group__"),
         .variable = str_replace(.variable, "__" , " - "),
         .variable = clean_varnames(.variable)) %>% 
  ggplot(aes(.value, .variable)) +
  stat_interval(.width = c(0.95, 0.89, 0.67, 0.50)) +
    geom_vline(xintercept = 0, linetype = "dashed") +
  labs(x = "Estimate", y = "Group", colour = "HDI (%)") +
  scale_colour_brewer(palette = "Oranges") +
  theme_classic() +
  theme(legend.position = "top",
        panel.grid.major.x = element_line(colour = "grey", linetype = "dotted"),
        axis.title.y = element_blank(),
        axis.text = element_text(colour = "black"))

@@

### 2020-08-06 Manuscript/te_manuscript.Rmd
Prepare report for meeting

Each trial started with a yellow fixation dot presented during one second on the center of the screen over a black background. After one second, the audio started playing while the dot remained being displayed until the audio offset. Upon the offset of the fixation point and audio, participants were prompted to write their answer by a ">" symbol. Typed letters were displayed in the screen in real time to provide visual feed-back to participants. Participants were allowed to correct their answer. Then, participants pressed the RETURN key to start and new trial. We excluded trials where participants did not type an existing word in the correspondent language, or did not type anything at all. Typos were allowed, however. Participants contributed a total of `r distinct(dat, participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` valid trials (`r filter(dat, test_language=="Catalan") %>%  distinct(participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` in Catalan, `r filter(dat, test_language=="Spanish") %>%  distinct(participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` in Spanish). The task took approximately 15 minutes to be completed.

@@

We divided participants in three groups: English natives tested in Catalan (ENG-CAT), English natives tested in Spanish (ENG-SPA), and Spanish natives tested in Catalan (SPA-CAT). We analysed the accuracy of responses to each item by calculating the proportion of correct responses across participants, resulting in one data point per item and group. We compared how our similarity score fitted the data in comparison with other measures:total shared phonemes, shared phonemic onset, and orthographical Damerau-Levenshtein distance. We fit several Bayesian linear models [@sorensen2015], each including one of the aforementioned variables as predictor, and the proportion of correct reponses as output. Continuous predictors were standadised and categorical predictors were sum-coded [@schad2020]. Missing predictor scores were imputed before fitting the models using multiple imputation [@vanbuuren2013]. To test and account for cross-group differences, we included a random intercept for each group. We compared models using Paretho-Smoothed Importance Sampling (PSIS-LOO, @vehtari2016). More information about the models and model comparison can be found in Appendix X.

>>>>

Each trial started with a yellow fixation dot presented during one second on the centre of the screen over a black background. After one second, the audio started playing while the dot remained being displayed until the audio offset. Upon the offset of the fixation point and audio, participants were prompted to write their answer by a ">" symbol. Typed letters were displayed in the screen in real time to provide visual feed-back to participants. Participants were allowed to correct their answer. Then, participants pressed the RETURN key to start and new trial. We excluded trials where participants did not type an existing word in the correspondent language, or did not type anything at all. Typos were allowed, however. Participants contributed a total of `r distinct(dat, participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` valid trials (`r filter(dat, test_language=="Catalan") %>%  distinct(participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` in Catalan, `r filter(dat, test_language=="Spanish") %>%  distinct(participant, trial_id) %>% nrow() %>% printnum(., big.mark = ",")` in Spanish). The task took approximately 15 minutes to be completed.

@@

We divided participants in three groups: English natives tested in Catalan (ENG-CAT), English natives tested in Spanish (ENG-SPA), and Spanish natives tested in Catalan (SPA-CAT). We analysed the accuracy of responses to each item by calculating the proportion of correct responses across participants, resulting in one data point per item and group. We compared how our similarity score fitted the data in comparison with other measures:total shared phonemes, shared phonemic onset, and orthographical Damerau-Levenshtein distance. We fit several Bayesian linear models [@sorensen2015], each including one of the aforementioned variables as predictor, and the proportion of correct responses as output. Continuous predictors were standardised and categorical predictors were sum-coded [@schad2020]. Missing predictor scores were imputed before fitting the models using multiple imputation [@vanbuuren2013]. To test and account for cross-group differences, we included a random intercept for each group. We compared models using Paretho-Smoothed Importance Sampling (PSIS-LOO, @vehtari2016). More information about the models and model comparison can be found in Appendix X.
